---
title: "Fig_3_a_b"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(scater)
library(RColorBrewer)
library(pheatmap)

#load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))

#mapping stromal cell types onto sce
Classes <- c("LC1", "LC2", "LP","BA", "HR", "EN",  "VA","IM", "FB")
Stromal <- c("Epithelial", "Epithelial", "Epithelial","Epithelial", "Epithelial", "Endothelial",  "Vascular_accessory","Immune", "Fibroblast")
colData(sce)$Stromal.clusters <- mapvalues(colData(sce)$Identity, Classes, Stromal)

pD <- as.data.frame(colData(sce))
```
*Check that the data is loaded correctly*
```{r}
sce
```
***Figure 3a***
*Stromal cells were classified into vascular accessory (VA), endothelial (EN), fibroblasts (FB) and immune (IM) cells.*
```{r}
#load stromal colours
Stromal <- c("Epithelial", "Endothelial", "Vascular_accessory","Immune", "Fibroblast")
stroma.colours <- c("gray90", "lightgoldenrod", "sandybrown", "#99A800", "tan3")

UMAP <- ggcells(sce, mapping=aes(x=UMAP1, y=UMAP2, color=Stromal.clusters)) + 
    geom_point() + scale_color_manual(values=setNames(stroma.colours, Stromal)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
UMAP

pdf("../../data/Fig_3/UMAP_Classes.stromal.pdf", width=12, height=8)
UMAP
dev.off()
```
***Figure 3b***
*Canonical stromal markers were used to classify the different stromal subtypes where LMC only contained IM cells.*
```{r}
set.seed(300)

#Load heatmap cluster colours
my_colour = list(
    Sample=c(HMC1 = "salmon1", HMC2 = "peachpuff2", HMC3 = "burlywood", HMC4 = "#F8766D", HMC2B="peachpuff3", HMC5="salmon2", HMC6="pink1", HMC7="indianred2", HMC8="indianred4", HMC9="hotpink2", RB1 = "steelblue3", RB2 = "slategray3", RB3 = "paleturquoise3", RB4 = "#00BFC4", RB5 = "turquoise", RB6 = "turquoise4", RB7 = "lightblue1", RB8 = "lightskyblue2" ),
    Identity=c(LC1= "slateblue2", LC2="plum", EN="lightgoldenrod", VA="sandybrown",LP="violetred3",BA="tomato1", HR="palevioletred3", IM="#99A800",FB="tan3"))

# Read in stromal gene list
genes <- c("SPARCL1","APOE","PDK4","ANGPT2","TFPI","AKAP12","MT1A","CDH5","CD93","CLDN5","ECSCR","ENG","S1PR1","SELE","TNFAIP6","MMP2","LUM","PDGFRA","COL1A1","IGFBP6","SERPINE2","CXCR4","STK4","CORO1A","CCL4","CD38","CD24","CD74")
 
# Subset stromal/immune cells
pD.sub <- subset(pD,pD$Stromal.clusters=="Endothelial"| pD$Stromal.clusters=="Vascular_accessory" | pD$Stromal.clusters=="Fibroblast" | pD$Stromal.clusters=="Immune")

subsP <- group_by(pD.sub, Identity) %>%
        mutate(Identity=forcats::fct_relevel(Identity, c("VA","EN","FB", "IM"))) %>%
        arrange(Identity, State)
subsP <- as.data.frame(subsP)

genes[(genes %in% rownames(logcounts(sce)))=="FALSE"]
mheat <- logcounts(sce)[genes,as.character(subsP$Barcode)]
mheat <- mheat/apply(mheat,1,max) # Scale to 0-1 for visualization

#Set data for heat-map column name annotations
annoCol <- data.frame(Sample=subsP$Sample, Identity=subsP$Identity)
row.names(annoCol) <- subsP$Barcode
annoCol %>% arrange(Identity, Sample)

# Plot heatmap
stro <-  pheatmap(mheat,
	 cluster_rows=FALSE,
	 cluster_cols=FALSE,
         show_rownames=TRUE,
         show_colnames=FALSE,
         annotation_legend=TRUE,
	       annotation_col=annoCol,
         annotation_colors=my_colour,
	       fontsize=8)

ggsave(filename="../../data/Fig_3/Heatmap_stromal_ClusterSignatures.pdf",stro[[4]],width=5,height=3.5)
```
*Saving session info*
```{r}
sessionInfo()
```