---
title: "Fig_4_a"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(knitr)

#libraries for reformatting data for bulk DEG analysis
library(SingleCellExperiment)
library(Seurat)

#load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
```
*Check that the data is loaded correctly*
```{r}
sce
```
*Setting colour charts*
```{r}
sample.set <- c("HMC1", "HMC2", "HMC3","HMC4", "HMC2B","HMC5","HMC6","HMC7","HMC8", "HMC9","RB1", "RB2","RB3","RB4","RB5","RB6","RB7", "RB8")
colour.names <- c("salmon1", "peachpuff2", "burlywood", "#F8766D","peachpuff3", "salmon2", "pink1",  "indianred2", "indianred4", "hotpink2", "steelblue3", "slategray3", "paleturquoise3","#00BFC4","turquoise", "turquoise4", "lightblue1", "lightskyblue2")
```
**Comparing human mammary signatures from Pfefferle et al. 2015**
*First generate the Seurat object so that the signatures can be compared across the cells*
```{r}
pD <- as.data.frame(colData(sce))

#Order cells by identity in the data frame
pD.ordered <- group_by(pD, Identity) %>%
  ungroup(Identity) %>%
        mutate(Identity=forcats::fct_relevel(Identity, c("LC1", "LC2", "LP","BA", "HR", "EN", "VA","IM", "FB"))) %>%
        arrange(Identity)

#Now reorganise the sce object
sce <- sce[,as.character(pD.ordered$Barcode)]

#Create a Seurat object
mammary <- Seurat::CreateSeuratObject(counts = logcounts(sce), min.cells = 3, project = "Mammary_cells")
mammary <- Seurat::AddMetaData(object = mammary, metadata = pD.ordered$Sample, col.name = "Sample")
mammary <- Seurat::AddMetaData(object = mammary, metadata = pD.ordered$Identity, col.name = "Identity")
head(mammary@meta.data)
```
*Importing mammary signatures from Pfefferle et al. 2015 and then mapping the signatures*
```{r}
#import published gene list which is taken from Table S1 from Pfefferle et al. 2015-> data is in form of Entrez ID
geneList1 <- read.csv("../../data/MammarySignatures_Pfefferle2015.csv",stringsAsFactors = FALSE)

#load annotation tool biomaRt
library("biomaRt")                                                                                                                   
listMarts()                                                                                                                           
ensembl <- useMart("ensembl", host="uswest.ensembl.org",dataset="hsapiens_gene_ensembl")                
filters = listFilters(ensembl)

#Create gene list for Basal cells
BC <- as.vector(geneList1$aMaSC.HsEnriched[3:129])
BC_pub <- as.data.frame(getBM(filters="entrezgene_id", attributes=c("entrezgene_id","hgnc_symbol"), values=BC, mart=ensembl))
length(BC) #127
dim(BC_pub) #128 2
setdiff(BC, BC_pub$entrezgene_id) #"10638" according to NCBI this gene is "SPHAR" but is a discontinued gene so will be removed from analysis
BC_list <- BC_pub
write.csv(BC_list, file= "../../data/Fig_4/RB_Published_Basal_Signature.csv")

#Create gene list for Luminal Progenitors
LP <- as.vector(geneList1$LumProg.HsEnriched[3:170])
LP_pub <- as.data.frame(getBM(filters="entrezgene_id", attributes=c("entrezgene_id","hgnc_symbol"), values=LP, mart=ensembl))
length(LP) #168
dim(LP_pub) #166 2
setdiff(LP, LP_pub$entrezgene_id) #"283412" "728715" "347292" according to NCBI these genes are: "RPL29P12", "LOC728715" and "RPL36P14"
LP_list <- rbind(LP_pub, c("283412", "RPL29P12"), c("728715","LOC728715"), c("347292", "RPL36P14"))
write.csv(LP_list, file= "../../data/Fig_4/RB_Published_LP_Signature.csv")

#Create gene list for Mature Luminal cells
HR <- as.vector(geneList1$MatureLum.HsEnriched[3:165])
HR_pub <- as.data.frame(getBM(filters="entrezgene_id", attributes=c("entrezgene_id","hgnc_symbol"), values=HR, mart=ensembl))
length(HR) #163
dim(HR_pub) #162 2
setdiff(HR, HR_pub$entrezgene_id) #"11257" according to NCBI this gene is TP53
HR_list <- rbind(HR_pub, c("11257", "TP53"))
write.csv(HR_list, file= "../../data/Fig_4/RB_Published_HR_Signature.csv")

#Stromal cells
SC <- as.vector(geneList1$aStr.HsEnriched[3:386])
SC_pub <- as.data.frame(getBM(filters="entrezgene_id", attributes=c("entrezgene_id","hgnc_symbol"), values=SC, mart=ensembl))
length(SC) #384
dim(SC_pub) #384 2
SC_list <- SC_pub
write.csv(SC_list, file= "../../data/Fig_4/RB_Published_Stromal_Signature.csv")
```
*Generating Basal score for all cell clusters*
```{r}
mammary <- AddModuleScore(mammary, features = list(BC_list$hgnc_symbol), ctrl = 20, name= "PublishedBasal")
names(x=mammary[[]])
#[1] "orig.ident"        "nCount_RNA"        "nFeature_RNA"   "PublishedBasal1"  

head(x=mammary[[]])
#The following features are not present in the object: LGALS7B, ZP4, not searching for symbol synonyms

FeatureScatter(mammary, feature1="PublishedBasal1", feature2="nFeature_RNA", group.by= "Identity", cols= colours.classes)

BC.plot <- VlnPlot(object = mammary, features = c("PublishedBasal1"),  pt.size = 0.1, group.by= "Identity", cols= colours.classes) +
      geom_hline(yintercept=0, linetype="dashed", color="red")
BC.plot
```
*Generating Luminal progenitor score for all cell clusters*
```{r}
mammary <- AddModuleScore(mammary, features = list(LP_list$hgnc_symbol), ctrl = 20, name= "PublishedLP")
names(x=mammary[[]])
#[1] "orig.ident"         "nCount_RNA"         "nFeature_RNA"       "Sample"             "prcntMito"          "GraphClusters"     
#[7] "S.Score"            "G2M.Score"          "Phase"              "LuminalProgenitor1" "HormoneResponsive1" "PublishedBasal1"            
#[13] "PublishedLP1"   

head(x=mammary[[]])
#The following features are not present in the object: PRR4, DEFB4B, KRTAP6-3, RPL29P12, LOC728715, RPL36P14, not searching for symbol synonyms

FeatureScatter(mammary, feature1="PublishedLP1", feature2="nFeature_RNA", group.by= "Identity", cols= colours.classes)

LP.plot <- VlnPlot(object = mammary, features = c("PublishedLP1"),  pt.size = 0.1, group.by= "Identity", cols= colours.classes) +
      geom_hline(yintercept=0, linetype="dashed", color="red")
LP.plot
```
*Generating Hormone responsive score for all cell clusters*
```{r}
mammary <- AddModuleScore(mammary, features = list(HR_list$hgnc_symbol), ctrl = 20, name= "PublishedHR")
names(x=mammary[[]])
#[1] "orig.ident"         "nCount_RNA"         "nFeature_RNA"       "Sample"             "prcntMito"          "GraphClusters"     
#[7] "S.Score"            "G2M.Score"          "Phase"              "LuminalProgenitor1" "HormoneResponsive1" "PublishedBasal1"       
#[13] "PublishedLP1"       "PublishedHR1"  

head(x=mammary[[]])

FeatureScatter(mammary, feature1="PublishedHR1", feature2="nFeature_RNA", group.by= "Identity", cols= colours.classes)

HR.plot <- VlnPlot(object = mammary, features = c("PublishedHR1"),  pt.size = 0.1, group.by= "Identity", cols= colours.classes) +
      geom_hline(yintercept=0, linetype="dashed", color="red")
HR.plot
```
*Generating Stromal score for all cell clusters*
```{r}
mammary <- AddModuleScore(mammary, features = list(SC_list$hgnc_symbol), ctrl = 20, name= "PublishedSC")
names(x=mammary[[]])
#[1] "orig.ident"         "nCount_RNA"         "nFeature_RNA"       "Sample"             "prcntMito"          "GraphClusters"     
#[7] "S.Score"            "G2M.Score"          "Phase"              "LuminalProgenitor1" "HormoneResponsive1" "PublishedBasal1"       
#[13] "PublishedLP1"       "PublishedHR1"       "PublishedSC1"   

head(x=mammary[[]])

FeatureScatter(mammary, feature1="PublishedSC1", feature2="nFeature_RNA", group.by= "Identity", cols= colours.classes)

HR.plot <- VlnPlot(object = mammary, features = c("PublishedSC1"),  pt.size = 0.1, group.by= "Identity", cols= colours.classes) +
      geom_hline(yintercept=0, linetype="dashed", color="red")
HR.plot
```
*Saving the published scores together*
```{r}
head(mammary@meta.data)

metaData <- as.data.frame(mammary@meta.data)
published <- data.frame(barcode=rownames(metaData),BasalScore =metaData$PublishedBasal1)
published$LPScore =metaData$PublishedLP1
published$HRScore =metaData$PublishedHR1
published$StromScore =metaData$PublishedSC1
write.csv(published, file= "../../data/Published_mammary_cell_scores.csv")

pD.ordered$BasalScore <- metaData$PublishedBasal1
pD.ordered$LPScore =metaData$PublishedLP1
pD.ordered$HRScore =metaData$PublishedHR1
pD.ordered$StromScore =metaData$PublishedSC1
```
***Fig 4a***
*Violin plots of the mammary cell scores for either i) hormone responsive (HR, mature luminal), ii) luminal progenitor (LP) iii) stromal or iv) myoepithelial cells across the major cell clusters identified in this study.*

**Fig 4ai**
*Replotting HR scores*
```{r}
HR <- ggplot(pD.ordered, aes(x=Identity, y=HRScore, fill=Identity)) +
    geom_jitter(aes(color=Sample)) +
    geom_violin(draw_quantiles=0.5) +
    ylab("HR score") +
    ggtitle("Overall_HRScore") +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    scale_fill_manual(values=setNames(colours.classes,classes)) + scale_color_manual(values=setNames(colour.names,sample.set))
HR

pdf("../../data/Fig_4/Fig_4_a_i.pdf", width=10, height=6)
HR
dev.off()
```
**Fig 4aii**
*Replotting LP scores*
```{r}
LP <- ggplot(pD.ordered, aes(x=Identity, y=LPScore, fill=Identity)) +
    geom_jitter(aes(color=Sample)) +
    geom_violin(draw_quantiles=0.5) +
    ylab("LP score") +
    ggtitle("Overall_LPScore") +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    scale_fill_manual(values=setNames(colours.classes,classes)) + scale_color_manual(values=setNames(colour.names,sample.set))
LP

pdf("../../data/Fig_4/Fig_4_a_ii.pdf", width=10, height=6)
LP
dev.off()
```
**Fig 4a iii**
*Replotting Stromal scores*
```{r}
Strom <- ggplot(pD.ordered, aes(x=Identity, y=StromScore, fill=Identity)) +
    geom_jitter(aes(color=Sample)) +
    geom_violin(draw_quantiles=0.5) +
    ylab("Stromal score") +
    ggtitle("Overall_StromScore") +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    scale_fill_manual(values=setNames(colours.classes,classes)) + scale_color_manual(values=setNames(colour.names,sample.set))
Strom

pdf("../../data/Fig_4/Fig_4_a_iii.pdf", width=10, height=6)
Strom 
dev.off()
```
**Fig 4aiv**
*Replotting Basal scores*
```{r}
Basal <- ggplot(pD.ordered, aes(x=Identity, y=BasalScore, fill=Identity)) +
    geom_jitter(aes(color=Sample)) +
    geom_violin(draw_quantiles=0.5) +
    ylab("Basal score") +
    ggtitle("Overall_BasalScore") +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    scale_fill_manual(values=setNames(colours.classes,classes)) + scale_color_manual(values=setNames(colour.names,sample.set))
Basal
pdf("../../data/Fig_4/Fig_4_a_iv.pdf", width=10, height=6)
Basal
dev.off()
```
*Saving session info*
```{r}
sessionInfo()
```