---
title: "Fig_3c"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(knitr)

#libraries for reformatting data for bulk DEG analysis
library(SingleCellExperiment)
library(edgeR)
library(EnhancedVolcano)

#specific for GO term analysis
library(topGO)
#library(org.Hs.eg.db)

#specific for dendogram+heatmap
library(cowplot)
library(viridis)
library(RColorBrewer)
library(pheatmap)
library(gridExtra)
library(gridGraphics)

#load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
pD <- as.data.frame(colData(sce))
```
*Subset immune cells*
```{r}
sce.sub <- sce[,(colData(sce)$Identity=="IM")]
pD.sub <- as.data.frame(colData(sce.sub))
```
*Principle components analysis & subclustering of all immune cells*
Finding the highly variable genes for downstream PCA and clustering analysis
```{r}
set.seed(60)
dec.sce <- modelGeneVar(sce.sub)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)

chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)

sce.hvg.sub <- sce.sub[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.sub
altExpNames(sce.hvg.sub)
sce.hvg.sub
```
*Principle component analysis (PCA)*
```{r}
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub, dimred="corrected")
reducedDimNames(sce.hvg.sub)
plotReducedDim(sce.hvg.sub, dimred="PCA", colour_by="Sample")
```
*Uniform manifold approximation and projection (UMAP)*
```{r}
set.seed(60)
sce.hvg.sub <- runUMAP(sce.hvg.sub, dimred="PCA", min_dist=0.6, n_neighbors=18)
plotReducedDim(sce.hvg.sub, dimred="UMAP", colour_by="Sample")
```
*Clustering the cells*
```{r}
g <- buildSNNGraph(sce.hvg.sub, k=15, use.dimred = "PCA")
clust <- igraph::cluster_louvain(g)$membership
table(clust)
colData(sce.hvg.sub)$GraphClusters.sub <- factor(paste0("C",clust))
colData(sce.sub)$GraphClusters.sub <- colData(sce.hvg.sub)$GraphClusters.sub

plotReducedDim(sce.hvg.sub, dimred="UMAP", colour_by="GraphClusters.sub")
plotReducedDim(sce.hvg.sub, dimred="UMAP", colour_by="GraphClusters")
```
*Adding PCA, UMAP and GraphClusterd to pD for ggplots visualisation*
```{r}
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
UMAP <- as.data.frame(reducedDim(sce.hvg.sub, "UMAP"))
pD.sub$GraphClusters.sub <- colData(sce.hvg.sub)$GraphClusters.sub

pD.sub$PCA1.sub <- PCA$PC1
pD.sub$PCA2.sub <- PCA$PC2
pD.sub$PCA3.sub <- PCA$PC3
pD.sub$PCA4.sub <- PCA$PC4
pD.sub$UMAP1.sub <- UMAP$V1
pD.sub$UMAP2.sub <- UMAP$V2
colData(sce.sub)$UMAP1.sub <- UMAP$V1
colData(sce.sub)$UMAP2.sub <- UMAP$V2

saveRDS(sce.hvg.sub,file="../../data/sce_immunesub.rds")
```
**Fig 3c**
*UMAP coloured by samples*
```{r}
#Sample colours
sample.set <- c("HMC1", "HMC2", "HMC3","HMC4", "HMC2B","HMC5","HMC6","HMC7","HMC8", "HMC9","RB1", "RB2","RB3","RB4","RB5","RB6","RB7", "RB8")
colour.names <- c("salmon1", "peachpuff2", "burlywood", "#F8766D","peachpuff3", "salmon2", "pink1",  "indianred2", "indianred4", "hotpink2", "steelblue3", "slategray3", "paleturquoise3","#00BFC4","turquoise", "turquoise4", "lightblue1", "lightskyblue2")

#Generate UMAP
UMAP.sample <- ggplot(pD.sub, aes(x=UMAP1, y=UMAP2, color=Sample))+
    geom_point() +
    xlab("UMAP1") +
    ylab("UMAP2") +
    ggtitle("UMAPs") +
    xlim(-5,5) +
    ylim(5,17) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colour.names, sample.set))
UMAP.sample 

pdf("../../data/Fig_3/Fig_3c_UMAP_Sample.pdf", width=6, height=4)
UMAP.sample
dev.off()
```
*UMAP coloured by sub clusters*
```{r}
#GraphClusters.sub colours
cluster.set.sub <- c("C1", "C2", "C3","C4", "C5", "C6","C7","C8", "C9","C10","C11", "C12", "C13","C14", "C15") 
colours.sub <- c("lightgoldenrod3", "lightgoldenrod4", "olivedrab", "palegreen3","lemonchiffon2", "gold4", "#99A800","gold", "darkolivegreen",  "lightsalmon3","yellow4", "tan1", "palegreen4", "palegreen3", "orange1")

#Generate UMAP
UMAP.subcluster <- ggplot(pD.sub, aes(x=UMAP1, y=UMAP2, color=GraphClusters.sub))+
    geom_point() +
    xlab("UMAP1") +
    ylab("UMAP2") +
    ggtitle("UMAPs") +
    xlim(-5,5) +
    ylim(5,17) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.sub, cluster.set.sub))
UMAP.subcluster

pdf("../../data/Fig_3/Fig_3c_UMAP_GraphClusters.sub.pdf", width=6, height=4)
UMAP.subcluster
dev.off()
```
*Umap coloured by gene*
```{r}
#Load colours
colour <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100)

#Convert data
m.imm <- as.matrix(logcounts(sce.sub))

#Load gene list
Genes <- c("CD3D", "JCHAIN", "CD68", "CD79A", "C1QB", "CD8A", "EOMES", "ICOS", "RORA", "FOXP3")

myGraphs <- list()
pdf("../../data/Fig_3/Fig_3c_UMAP_genes.pdf", width=6, height=5)
for (Gene in Genes) {
myGene <- as.numeric(m.imm[Gene,])

umap.gene <- ggcells(sce.sub, aes(x=UMAP1.sub, y=UMAP2.sub, color=myGene))+
    scale_color_gradientn(colors=colour) +
    geom_point() +
  #  xlab("UMAP 1") +
  #  ylab("UMAP 2") +
    ggtitle(Gene) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_blank(), legend.position = c(0.9,0.9), legend.title = element_blank(), plot.title = element_text(family="sans", face="bold.italic", size=45, vjust=-2, hjust=0.05), axis.text.x= element_blank(), axis.text.y = element_blank(), axis.title.x= element_blank(), axis.title.y= element_blank(), axis.ticks = element_blank())

myGraphs[[Gene]] <- umap.gene
print(myGraphs[[Gene]])
}
dev.off()
```
*Saving session info*
```{r}
sessionInfo()
```