---
title: "Fig_2_c"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(knitr)

#libraries for reformatting data for bulk DEG analysis
library(SingleCellExperiment)
library(edgeR)
library(EnhancedVolcano)

#load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))
```
*Check that the data is loaded correctly*
```{r}
sce
```
*Subsetting only luminal LMCs*
```{r}
set.seed(60)

sce.LumHMC <- sce[,colData(sce)$Identity=="LC1"|colData(sce)$Identity=="LC2"]
sce.sub <- sce.LumHMC[,colData(sce.LumHMC)$State=="HM"]
```
*Running DEG analysis*
Generating psuedo-bulk samples from luminal HMCs which have equal numbers of cells per sample
```{r}
#Create pseudo-bulk samples from the Samples subsetted according to GraphClusters
summed <- aggregateAcrossCells(sce.sub,
                               ids=DataFrame(label=sce.sub$Identity,
                                             sample=sce.sub$Sample))
summed
table(summed$sample,summed$label)
```
*DEG analysis on subsetted HMC luminal cells from LC1 and LC2 *
```{r}
y<- DGEList(counts(summed), samples = colData(summed))
y

#pre-processing
discarded <- isOutlier(y$samples$lib.size, log=TRUE, type="lower")
y <- y[,!discarded]
summary(discarded)

#remove lowly expressed genes
keep <- filterByExpr(y, group=summed$Identity)
y <- y[keep,]
summary(keep)

#correct for compositional bias
y <- calcNormFactors(y)
y$samples

#Stat modelling
design <- model.matrix(~factor(batch) + factor(Identity) , y$samples)
design

#estimate binomial (NB) dispersions
y <- estimateDisp(y,design)
summary(y$trended.dispersion)

#Plot
plotBCV(y)

#find fit
fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.prior)
summary(fit$df.prior)

#plot
plotQLDisp(fit)

#run test
res <- glmQLFTest(fit, coef=ncol(design))
summary(decideTests(res))

#Adding false discovery rate (FDR) values to the res$table and saving this as a new object DEGs with the rows in order of the FDR
topTags(res)
res$table$FDR <- p.adjust(res$table$PValue, method="BH")
DEGs <- res$table[order(res$table$FDR),]
dim(DEGs) #12703
write.csv(DEGs, file= "../../data/Supp_Tables/Supplementary_Table_1.csv")
```
***Fig 2c***
*Volcano plot displaying the findings of the differential gene expression analysis revealed 1,640 genes more highly expressed in LC1 compared to 1,782 genes highly expressed. Significant genes in red with the top 10 being annotated.*
```{r}
##Specifying labels for top 10 genes
sig.genes <- subset(DEGs, DEGs$FDR < 10e-8)

#Volcano plot
upgenes <- sig.genes[order(-sig.genes$logFC),]
uplab <- rownames(upgenes)[1:20]
downgenes <- sig.genes[order(sig.genes$logFC),]
downlab <- rownames(downgenes)[1:20]
labels <- c(uplab, downlab)

#Volcano plot
Volcano <- EnhancedVolcano(DEGs,
lab = rownames(DEGs),
x = 'logFC',
y = 'FDR',
title = 'DEGs HMC cluster 1 vs cluster 2',
subtitle = '',
subtitleLabSize = 2,
legendPosition = "bottom",
pointSize = 3.5,
labSize = 3.0,
FCcutoff = 0,
pCutoff = 10e-8,
col = c("grey", "grey", "steelblue", "red"),
selectLab = labels,
#legendVisible = FALSE,
drawConnectors = TRUE,
#boxedLabels = TRUE,
widthConnectors = 0.5,
legendIconSize = 4.0,
labFace = 'bold',
typeConnectors = 'open')
Volcano

pdf("../../data/Fig_2/Fig2c.pdf", width=12, height=8)
Volcano
dev.off()
```
*Saving session info*
```{r}
sessionInfo()
```