---
title: "Fig_3_d"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output: html_document
---
Commands taken from tutorial at: https://htmlpreview.github.io/?https://github.com/sqjin/CellChat/blob/master/tutorial/CellChat-vignette.html

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)

#Bioc Packages
library(scran)
library(scater)
library(CellChat)
library(patchwork)
library(RColorBrewer)

#Riverplots etc.
library(NMF)
library(ggalluvial)

#---#Read in annotated immune cells and subset for milk cells only#---#
sce.immune <- readRDS("../../data/RB_HMC_merged/sce_immunesub.rds")
sce.mini <- sce.immune[,colData(sce.immune)$State=="HM"]
sce.mini
#Add info about the clusters
clusters <- c("C1","C10","C3","C5","C4", "C6","C11", "C12","C9", "C8", "C7", "C2")
names <- c("Tcell","Tcell","Tcell","Tcell","Bcell_Plasma","Bcell_Plasma","Mono","Mono","Mono","Myeloid","Macro","Macro")
sce.mini$Ident.Detailed <- mapvalues(colData(sce.mini)$GraphClusters.sub, clusters, names)
pD.immune <- data.frame(Ident.Detailed =colData(sce.mini)$Ident.Detailed,row.names = colData(sce.mini)$Barcode)

#---#Read in all data#---#
sce <- readRDS("../../data/RB_HMC_merged/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/RB_HMC_merged/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))
sce.milk <- sce[,colData(sce)$State=="HM"]
sce.milk
#subsetting just epithelial cells
sce.epi <- sce.milk[,(colData(sce.milk)$Identity=="LC1"|colData(sce.milk)$Identity=="LC2")]
colData(sce.epi)$Ident.Detailed <- colData(sce.epi)$Identity
#subsetting just immune cells and annotating with info from the detailed analysis
sce.imm <- sce.milk[,colData(sce.milk)$Identity=="IM"]
colData(sce.imm) <- cbind(colData(sce.imm), pD.immune)

#joining the data back together with the correct annotations
sce.sub <- cbind(sce.epi,sce.imm)
```
*Check that data was read in and converted correctly*
```{r}
sce.sub
```
*Part I: Create the CellChat object*
```{r}
cellchat <- createCellChat(object = sce.sub, group.by = "Ident.Detailed")
cellchat <- setIdent(cellchat, ident.use = "Ident.Detailed") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
cellchat@idents = droplevels(cellchat@idents, exclude = setdiff(levels(cellchat@idents),unique(cellchat@idents))) #exclude unused labels
levels(cellchat@idents) # check factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group

#Set the ligand-receptor
CellChatDB <- CellChatDB.human
showDatabaseCategory(CellChatDB)
dplyr::glimpse(CellChatDB$interaction) #Show the structure of the database
CellChatDB.use <- CellChatDB # simply use the default CellChatDB
cellchat@DB <- CellChatDB.use # set the used database in the object

#Preprocessing expression data for cell-cell communication
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- projectData(cellchat, PPI.human) #project gene expression data onto PPI network (optional)
```
*Part II: Inference of cell-cell communication network*
Compute the communication probability and infer cellular communication network
```{r}
cellchat <- computeCommunProb(cellchat)
cellchat <- filterCommunication(cellchat, min.cells = 10) # Filter out the cell-cell communication if there are only few number of cells in certain cell groups
#The cell-cell communication related with the following cell groups are excluded due to the few number of cells:  Plasma

#exporting the ligand/receptors
df.net <- subsetCommunication(cellchat)
write.csv(df.net, file= "../../data/Supp_Tables/Supplementary_Table_5.csv")

#Infer the cell-cell communication at a signaling pathway level
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat) #Calculate the aggregated cell-cell communication network

#Save all signalling pathways showing significant communications
Pthways <- cellchat@netP$pathways
write.csv(Pthways, file= "../../data/CellChat_AllSigPathways.csv")

#Save the CellChat object
saveRDS(cellchat, file = "../../data/CellChat.milkCells.rds")
```
*Automatically save the plots of the all inferred network for quick exploration*
Be aware that this will generate a lot of plots (all the significant ligand-receptor plots), however it is very useful in identifying pathways with lots of receptor-ligand pairs across many different cell subtypes.
```{r}
# Access all the signaling pathways showing significant communications
pathways.show.all <- cellchat@netP$pathways
# check the order of cell identity to set suitable vertex.receiver
levels(cellchat@idents)
vertex.receiver = seq(1,4)
for (i in 1:length(pathways.show.all)) {
  # Visualize communication network associated with both signaling pathway and individual L-R pairs
  netVisual(cellchat, signaling = pathways.show.all[i], vertex.receiver = vertex.receiver, layout = "hierarchy")
  # Compute and visualize the contribution of each ligand-receptor pair to the overall signaling pathway
  gg <- netAnalysis_contribution(cellchat, signaling = pathways.show.all[i])
  ggsave(filename=paste0(pathways.show.all[i], "_L-R_contribution.pdf"), plot=gg, width = 3, height = 2, units = 'in', dpi = 300)
}
```
***Fig 3dii***
*Observed interactions between the cell types.*
```{r}
#groupSize <- as.numeric(table(cellchat@idents))
#par(mfrow = c(1,2), xpd=TRUE)

#plot interaction map
num <- netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions", color.use = c("slateblue2", "plum", "lightsalmon3", "darkolivegreen", "#99A800", "gold4", "gold"),vertex.label.cex = 2)
num
pdf("../../data/Fig_3/Fig3dii.pdf", width=12, height=8)
num
dev.off()
```
***Fig 3diii***
*All immune cell subtypes signal to LC1 via the EGF, MK and SPP1 signalling pathway (selected receptor-ligand pairs shown). Whereas LC1 and LC2 luminal clusters from milk are signalling to milk immune cells via MHC-II, MHC-I, CSF and GRN signalling pathways.*
```{r}
#EGF
pathways.show <- "EGF"
pairLR.EGF <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.EGF[5,] #HBEGF-EGFR
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
#MK
pathways.show <- "MK"
pairLR.MK <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MK[6,] #MDK-NCL
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
#SPP1
pathways.show <- "SPP1"
pairLR.SPP1 <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.SPP1[1,] #SPP1-CD44
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
LR.show <- pairLR.SPP1[3,] #SPP1-(ITAV+ITGB6)
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
#MHC-II
pathways.show <- "MHC-II"
pairLR.MHCII <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHCII[4,] #HLC-DMA-CD4
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
LR.show <- pairLR.MHCII[8,] #HLA-DRA-CD4
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
#MHC-I
pathways.show <- "MHC-I"
pairLR.MHCI <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHCI[1,] #HLA-F-LILRB1
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
#GRN
pathways.show <- "GRN"
pairLR.GRN <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.GRN[1,] #GRN-SORT1
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
#CSF
pathways.show <- "CSF"
pairLR.CSF <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.CSF[1,] #IL34-CSF1R
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle", color.use= color.use)
```
*Return the session information*
```{r}
sessionInfo()
```

