---
title: "B2_Downstream_analysis"
author: "Alecia"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
Using our cleaned data sce_clean.rds we aim to understand more about the structure of the data that we are working with so that we can try and interpret the biological meaning it might have.

**Setting up**
*Load data and packages*
Firstly we must load the important libraries and then the data which is in a SingleCellExperiment format.
```{r, message=FALSE}
#load libraries
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(Rtsne)
library(dynamicTreeCut)
library(igraph)
library(Matrix)
library(umap)
library(PCAtools)

##load data
sce <- readRDS("../../data/Batch2/sce_clean.rds")
```
*Check that the data is loaded correctly*
```{r}
sce
```
**Determining highly variable genes (hvg)**
*Mean expression and variance of each gene*
Here we can determine the mean expression and variance for each gene and generate a trend line that fits this relationship. If a gene fits along this trend line, this suggests that the gene is stably expressed and is not really of interest for down stream clustering and cell identification.
```{r}
dec.sce <- modelGeneVar(sce)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
```
*Selecting highly variable genes*
Next we select the highly variable genes as the top 10% of genes and store them as an object "chosen".
```{r}
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
```
To store this object, we subset the original "sce" object by the highly variable genes and create a new object called "sce.hvg". Next we store the old "original" sce object (containing original colData and counts matrix) is now stored in the new "sce.hvg" object.
```{r}
sce.hvg <- sce[chosen,]
dim(sce.hvg)

altExp(sce.hvg, "original") <-sce
altExpNames(sce.hvg)
sce.hvg
```
**Principle component analysis**
```{r}
set.seed(100)
sce.hvg <- runPCA(sce.hvg)
reducedDimNames(sce.hvg)

plotReducedDim(sce.hvg, dimred="PCA", colour_by="Sample")
```
**Uniform manifold approximation and projection**
```{r}
set.seed(60)
sce.hvg <- runUMAP(sce.hvg, dimred="PCA", min_dist=0.7, n_neighbors=18)
plotReducedDim(sce.hvg, dimred="UMAP", colour_by="Sample")
```
**Clustering the cells**
```{r}
g <- buildSNNGraph(sce.hvg, k=30, use.dimred = "PCA")
clust <- igraph::cluster_louvain(g)$membership
table(clust)
colData(sce.hvg)$GraphClusters <- factor(paste0("C",clust))
```

```{r}
plotReducedDim(sce.hvg, dimred="UMAP", colour_by="GraphClusters")
```
*Ensuring that both the sce.hvg subset and full sce plot are stored*
```{r}
sce.original <- altExp(sce.hvg, "original", withColData=TRUE)
reducedDim(sce.original, "PCA") <- reducedDim(sce.hvg, "PCA")
reducedDim(sce.original, "UMAP") <- reducedDim(sce.hvg, "UMAP")
altExp(sce.original, "hvg") <- sce.hvg

#Adding PCA and UMAP data to colData
PCA <- as.data.frame(reducedDim(sce.original, "PCA"))
UMAP <- as.data.frame(reducedDim(sce.original, "UMAP"))
colData(sce.original)$PCA1 <- PCA$PC1
colData(sce.original)$PCA2 <- PCA$PC2
colData(sce.original)$PCA3 <- PCA$PC3
colData(sce.original)$PCA4 <- PCA$PC4
colData(sce.original)$UMAP1 <- UMAP$V1
colData(sce.original)$UMAP2 <- UMAP$V2

sce.original
```
*Making some extra plots to examine the UMI/gene content of cells*
```{r}
pD <-as.data.frame(colData(sce.original))

UMI <- ggplot(pD, aes(x=GraphClusters, y=sum, fill=GraphClusters)) +
    geom_jitter(aes(color=Sample)) +
    geom_violin(draw_quantiles=0.5) +
    ylab("UMI") +
    ggtitle("UMI") +
    scale_y_log10() +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
UMI

pdf("../../data/Batch2/Downstream_graphs/UMI_scores.pdf", width=10, height=6)
UMI
dev.off()

Genes <- ggplot(pD, aes(x=GraphClusters, y=detected, fill=GraphClusters)) +
    geom_jitter(aes(color=Sample)) +
    geom_violin(draw_quantiles=0.5) +
    ylab("Genes") +
    ggtitle("Genes") +
    scale_y_log10() +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
Genes

pdf("../../data/Batch2/Downstream_graphs/Genes_scores.pdf", width=10, height=6)
Genes
dev.off()


UMAP <- ggcells(sce.original, mapping=aes(x=UMAP1, y=UMAP2, color=sum)) + 
    geom_point() +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
UMAP

pdf("../../data/Batch2/Downstream_graphs/UMAP_UMI.pdf", width=12, height=8)
UMAP
dev.off()

UMAP <- ggcells(sce.original, mapping=aes(x=UMAP1, y=UMAP2, color=detected)) + 
    geom_point() +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
UMAP

pdf("../../data/Batch2/Downstream_graphs/UMAP_genes.pdf", width=12, height=8)
UMAP
dev.off()
```
*Saving the data*
```{r}
saveRDS(sce.original,file="../../data/Batch2/sce_final.rds")
sessionInfo()
```
