---
title: "B1_Downstream_analysis"
author: "Alecia"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
Using our cleaned data sce_clean.rds we aim to understand more about the structure of the data that we are working with so that we can try and interpret the biological meaning it might have.

**Setting up**
*Load data and packages*
Firstly we must load the important libraries and then the data which is in a SingleCellExperiment format.
```{r, message=FALSE}
#load libraries
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(Rtsne)
library(dynamicTreeCut)
library(igraph)
library(Matrix)
library(umap)
library(PCAtools)

##load data
sce <- readRDS("../../data/Batch1/sce_clean.rds")
```
*Check that the data is loaded correctly*
```{r}
sce
```
**Determining highly variable genes (hvg)**
*Mean expression and variance of each gene*
Here we can determine the mean expression and variance for each gene and generate a trend line that fits this relationship. If a gene fits along this trend line, this suggests that the gene is stably expressed and is not really of interest for down stream clustering and cell identification.
```{r}
dec.sce <- modelGeneVar(sce)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
```
*Selecting highly variable genes*
Next we select the highly variable genes as the top 10% of genes and store them as an object "chosen".
```{r}
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
```
To store this object, we subset the original "sce" object by the highly variable genes and create a new object called "sce.hvg". Next we store the old "original" sce object (containing original colData and counts matrix) is now stored in the new "sce.hvg" object.
```{r}
sce.hvg <- sce[chosen,]
dim(sce.hvg)

altExp(sce.hvg, "original") <-sce
altExpNames(sce.hvg)
sce.hvg
```
**Principle component analysis**

```{r}
set.seed(100)
sce.hvg <- runPCA(sce.hvg)
reducedDimNames(sce.hvg)

plotReducedDim(sce.hvg, dimred="PCA", colour_by="Sample")
```
Finding the elbow point for the data
```{r}
percent.var <- attr(reducedDim(sce.hvg,"PCA"), "percentVar")
chosen.elbow <- PCAtools::findElbowPoint(percent.var)
chosen.elbow

plot(percent.var, xlab="PC", ylab="Variance explained (%)")
abline(v=chosen.elbow, col="red")
```
Here we are trying to figure out the number of PCs (d) that should be kept for downstream analysis. To do this we are comparing the "d" to the number of clusters generated using getClusteredPCs(pcs). Here it appears that d should equal 50.

```{r}
pcs <- reducedDim(sce.hvg, "PCA")
choices <- getClusteredPCs(pcs)
metadata(choices)$chosen

plot(choices$n.pcs, choices$n.clusters, xlab="Number of PCs", ylab="Number of clusters")
abline(a=1, b=1, col="red")
abline(v=metadata(choices)$chosen, col="grey80", lty=2)
```
The results from the above test suggest that we only require 47 out of the total 50 PC's and therefore we subset the data to reflect this.
```{r}
reducedDim(sce.hvg, "PCA") <- reducedDim(sce.hvg, "PCA")[,1:47]
ncol(reducedDim(sce.hvg, "PCA"))
```
**Uniform manifold approximation and projection**
```{r}
set.seed(60)
sce.hvg <- runUMAP(sce.hvg, dimred="PCA", min_dist=0.7, n_neighbors=18)
plotReducedDim(sce.hvg, dimred="UMAP", colour_by="Sample")
```
**Clustering the cells**
```{r}
g <- buildSNNGraph(sce.hvg, k=30, use.dimred = "PCA")
clust <- igraph::cluster_louvain(g)$membership
table(clust)
colData(sce.hvg)$GraphClusters <- factor(paste0("C",clust))
```

```{r}
plotReducedDim(sce.hvg, dimred="UMAP", colour_by="GraphClusters")
```
*Ensuring that both the sce.hvg subset and full sce plot are stored*
```{r}
sce.original <- altExp(sce.hvg, "original", withColData=TRUE)
reducedDim(sce.original, "PCA") <- reducedDim(sce.hvg, "PCA")
reducedDim(sce.original, "UMAP") <- reducedDim(sce.hvg, "UMAP")
altExp(sce.original, "hvg") <- sce.hvg

#Adding PCA and UMAP data to colData
PCA <- as.data.frame(reducedDim(sce.original, "PCA"))
UMAP <- as.data.frame(reducedDim(sce.original, "UMAP"))
colData(sce.original)$PCA1 <- PCA$PC1
colData(sce.original)$PCA2 <- PCA$PC2
colData(sce.original)$PCA3 <- PCA$PC3
colData(sce.original)$PCA4 <- PCA$PC4
colData(sce.original)$UMAP1 <- UMAP$V1
colData(sce.original)$UMAP2 <- UMAP$V2

sce.original
```
**Saving the data**
```{r}
cDat <- logcounts(sce.original)
pDat <- colData(sce.original)
fDat <- rowData(sce.original)
DataList <- list("phenoData"=pDat, "featureData"=fDat, "counts"=cDat)
saveRDS(DataList,file="../../data/Batch1/ExpressionList_final.rds")

saveRDS(sce.original,file="../../data/Batch1/sce_final.rds")
sessionInfo()
```
