---
title: "Supp_Fig_14"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
Before running this analysis, make sure that "Fig_3_d.Rmd" has been run so that the "CellChat.milkCells.rds" is available
```{r, message=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)

#Bioc Packages
library(scran)
library(scater)
library(CellChat)
library(patchwork)
library(RColorBrewer)

#Riverplots etc.
library(NMF)
library(ggalluvial)

#---#Read in annotated immune cell chat object#---#
cellchat <- readRDS("../../data/RB_HMC_merged/CellChat/CellChat.milkCells.rds")
```
***Supplementary Fig 14a***
*Chord plots displaying an overview of all receptor-ligand signalling occurring between luminal and immune milk cells in the pathways of EGF, MK, SPP1, MHC-II, MHC-I, GRN and CSF.*
```{r}
color.use <- c("slateblue2", "plum", "lightsalmon3", "darkolivegreen", "#99A800", "gold4", "gold")

netVisual_chord_gene(cellchat, sources.use = c(1:7), targets.use = c(1:7), signaling = c("EGF"),legend.pos.x = 8, color.use= color.use)
netVisual_chord_gene(cellchat, sources.use = c(1:7), targets.use = c(1:7), signaling = c("MK"),legend.pos.x = 8, color.use= color.use)
netVisual_chord_gene(cellchat, sources.use = c(1:7), targets.use = c(1:7), signaling = c("SPP1"),legend.pos.x = 8, color.use= color.use)
netVisual_chord_gene(cellchat, sources.use = c(1:2), targets.use = c(1:7), signaling = c("MHC-II"),legend.pos.x = 8, color.use= color.use)
netVisual_chord_gene(cellchat, sources.use = c(1:7), targets.use = c(1:7), signaling = c("MHC-I"),legend.pos.x = 8, color.use= color.use)
netVisual_chord_gene(cellchat, sources.use = c(1:7), targets.use = c(1:7), signaling = c("CSF"),legend.pos.x = 8, color.use= color.use)
netVisual_chord_gene(cellchat, sources.use = c(1:7), targets.use = c(1:7), signaling = c("GRN"),legend.pos.x = 8, color.use= color.use)
```
***Supplementary Fig 14b***
*Gene expression from all milk cell subtypes for each highlighted pathway. Arrows indicate that these genes are involved in both MK and SPP1 pathways.*
```{r}
pD <- as.data.frame(colData(sce.sub))
m <- logcounts(sce.sub)

MHCII <- c("HLA-DPA1", "HLA-DPB1", "HLA-DQA1", "HLA-DMA", "HLA-DMB", "HLA-DOA", "HLA-DQB1", "HLA-DRA", "HLA-DRB1","HLA-DRB5", "CD4")
MHCI <- c("HLA-F", "LILRB1", "IL34")
CSF <- c("CSF1", "CSF1R")
GRN <- c("GRN", "SORT1")
EGF <- c("EGF","AREG","HBEGF", "EREG", "EGFR", "ERBB2")
MK <- c("MDK", "SDC2", "SDC4", "LRP1", "NCL", "ITGA4", "ITGA6", "ITGB1")
SPP1 <- c("SPP1", "CD44", "ITGAV","ITGA4","ITGA5","ITGB1","ITGB6")
markers <- c(MHCII,MHCI,CSF,GRN,EGF,MK,SPP1)
grps <- colData(sce.sub)$Ident.Detailed

table(grps)

#Make bubblePlot function
bubblePlot <- function(m, markers, grps, cluster_col=FALSE,
cluster_row=FALSE, angled=TRUE, zscores=FALSE) {
  require(reshape2)
     out <- data.frame(numeric(length(markers)))
     colnames(out) <- levels(factor(grps))[1]
     out.freq <- out
     for (cond in levels(factor(grps))) {
     expr <- rowMeans(m[markers,grps==cond])
     expr.freq <- rowMeans(m[markers,grps==cond]>0)
     colname <- cond
     out[,colname] <- expr
     out.freq[,colname] <- expr.freq
     }
     out <- as.matrix(out)
     out.freq <- as.matrix(out.freq)
     rownames(out.freq) <- rownames(out) <- markers
     if(zscores) {
     out <- t(scale(t(out)))
     } else {
     out <- out/rowMax(out)
     }
     out.long <- melt(out,value.name="Mean")
     out.freq <- melt(out.freq,value.name="Frequency")
     out.long$Frequency <- out.freq$Frequency * 100

     if (cluster_row) {
     dis <- dist(out)
     hclst <- hclust(dis,method="ward.D2")
     lvlsVar1 <- rownames(out)[hclst$order]
     out.long$Var1 <- factor(out.long$Var1, levels=lvlsVar1)
     }
     if (cluster_col) {
     dis <- dist(t(out))
     hclst <- hclust(dis,method="ward.D2")
     lvlsVar2 <- colnames(out)[hclst$order]
     out.long$Var2 <- factor(out.long$Var2, levels=lvlsVar2)
     }
     p <- ggplot(out.long, aes(x=Var2, y=Var1, color=Mean,
size=Frequency)) +
     geom_point() +
     scale_color_gradient2(low="blue",high="red", mid="grey") +
    # scale_color_distiller(palette="Spectral") +
     scale_size(range=c(0,3)) +
theme(panel.grid.major=element_line(colour="grey80",size=0.1,linetype="dashed"),
       axis.text=element_text(size=7),
       legend.position="bottom",
       legend.direction="horizontal") +
     xlab("") +
     ylab("")
     if (angled) {
     p <- p + theme(axis.text.x=element_text(size=9, angle=45, hjust=1,
vjust=1))
     }
     return(p)
}

#now this is a normal ggplot that can be manipulated with gg-commands
Bubble.gene <- bubblePlot(m, markers, grps) + coord_flip() +# scale_x_reverse() +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 90), axis.line = element_line(colour = "black"))

pdf("../../data/Supp_Fig_14/Supp_Fig_14b.pdf", width=8, height=4.5)
Bubble.gene
dev.off()

```
*Return the session information*
```{r}
sessionInfo()
```

