---
title: "Supp_Fig_9"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(scater)
library(RColorBrewer)
library(pheatmap)

##load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))

pD <- as.data.frame(colData(sce))
```
*Check that the data is loaded correctly*
```{r}
sce
```
***Supp_Fig_9***
*Heatmap displaying the expression of key genes previously described in human milk cells across all lactation associated mammary cells (LMC) and non-lactation associated mammary cells (NMC) subtypes.*
```{r}
set.seed(300)

#Select colours
my_colour = list(
    Sample=c(HMC1 = "salmon1", HMC2 = "peachpuff2", HMC3 = "burlywood", HMC4 = "#F8766D", HMC2B="peachpuff3", HMC5="salmon2", HMC6="pink1", HMC7="indianred2", HMC8="indianred4", HMC9="hotpink2", RB1 = "steelblue3", RB2 = "slategray3", RB3 = "paleturquoise3", RB4 = "#00BFC4", RB5 = "turquoise", RB6 = "turquoise4", RB7 = "lightblue1", RB8 = "lightskyblue2" ),
    Identity=c(LC1= "slateblue2", LC2="plum", EN="lightgoldenrod", VA="sandybrown",LP="violetred3",BA="tomato1", HR="palevioletred3", IM="#99A800",FB="tan3"))

# Read in milk cell literature gene list
genes <- c("ABCG2","ATXN1","CD44","ENG","THY1","VIM","EPCAM","ITGA4","KRT14","KRT18","LALBA","POU5F1","SOX2","NANOG","KLF4","PODXL","NES","ITGB1","KIT","KRT5","PROM1")

# Sample cells from large clusters
pD.sub <- subset(pD, pD$Epithelial.clusters=="Luminal_HMC1"| pD$Epithelial.clusters=="Luminal_HMC2")

subsP <- group_by(pD, Identity) %>%
        do(sample_n(.,1000)) %>%
        mutate(Identity=forcats::fct_relevel(Identity, c("LC1", "LC2", "LP", "HR", "BA", "VA", "EN","FB","IM"))) %>%
        arrange(Identity, State)

subsP <- as.data.frame(subsP)

mheat <- as.matrix(logcounts(sce)[genes,as.character(subsP$Barcode)])
mheat <- mheat/apply(mheat,1,max) # Scale to 0-1 for visualization

#Set data for heat-map column name annotations
annoCol <- data.frame(Sample=subsP$Sample, Identity=subsP$Identity)
row.names(annoCol) <- subsP$Barcode
annoCol %>% arrange(Identity, Sample)

# Plot heatmap
h <-  pheatmap(mheat,
	 cluster_rows=FALSE,
	 cluster_cols=FALSE,
         show_rownames=TRUE,
         show_colnames=FALSE,
         annotation_legend=TRUE,
	       annotation_col=annoCol,
         annotation_colors=my_colour,
	       fontsize=12)

ggsave(filename="../../data/Supp_Fig_9/Heatmap_stem_ClusterSignatures.pdf",h[[4]],width=8,height=5)

```
*Saving session info*
```{r}
sessionInfo()
```

