---
title: "Supp_Fig_8"
author: "Alecia"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(knitr)

#load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))
```
*Check that the data is loaded correctly*
```{r}
sce
```
*Setting colourS*
```{r}
#UMAP cluster colours
cluster.set <- ClustCol$Clusters
colours <- ClustCol$Clusters.colour
```
**Supp_Fig_8**
*Principal component analysis of individual milk donor samples reveals cells separate into luminal clusters LC1 and LC2.*


*Subsetting by Sample first then running PCA*
*LMC1*
```{r}
#LMC1/HMC1
set.seed(60)
#subsetting on just HMC1
sce.HMC1 <- sce[,colData(sce)$Sample=="HMC1"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC1)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC1[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC1
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../data/Supp_Fig_8/PCA_1_2_HMC1_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC2*
```{r}
#HMC2
set.seed(60)
#subsetting on just HMC2
sce.HMC2 <- sce[,colData(sce)$Sample=="HMC2"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC2)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC2[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC2
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../data/Supp_Fig_8/PCA_1_2_HMC2_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC2B*
```{r}
#HMC2B
set.seed(60)
#subsetting on just HMC2B
sce.HMC2B <- sce[,colData(sce)$Sample=="HMC2B"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC2B)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC2B[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC2B
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA and colour
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../data/Supp_Fig_8/PCA_1_2_HMC2B_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC3*
```{r}
#HMC3
set.seed(60)
#subsetting on just HMC3
sce.HMC3 <- sce[,colData(sce)$Sample=="HMC3"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC3)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC3[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC3
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../data/Supp_Fig_8/PCA_1_2_HMC3_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC4*
```{r}
#HMC4
set.seed(60)
#subsetting on just HMC4
sce.HMC4 <- sce[,colData(sce)$Sample=="HMC4"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC4)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC4[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC4
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)
reducedDimNames(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../Supp_Fig_8/PCA_1_2_HMC4_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC5*
```{r}
#HMC5
set.seed(60)
#subsetting on just HMC5
sce.HMC5 <- sce[,colData(sce)$Sample=="HMC5"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC5)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC5[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC5
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA and colour
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../Supp_Fig_8/PCA_1_2_HMC5_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC6*
```{r}
#HMC6
set.seed(60)
#subsetting on just HMC6
sce.HMC6 <- sce[,colData(sce)$Sample=="HMC6"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC6)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC6[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC6
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../Supp_Fig_8/PCA_1_2_HMC6_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC7*
```{r}
#HMC7
set.seed(60)
#subsetting on just HMC7
sce.HMC7 <- sce[,colData(sce)$Sample=="HMC7"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC7)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC7[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC7
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA and colour
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../Supp_Fig_8/PCA_1_2_HMC7_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*LMC8*
```{r}
#HMC8
set.seed(60)
#subsetting on just HMC8
sce.HMC8 <- sce[,colData(sce)$Sample=="HMC8"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC8)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC8[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC8
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA and colour
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../data/Supp_Fig_8/PCA_1_2_HMC8_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*HMC9*
```{r}
#HMC9
set.seed(60)
#subsetting on just HMC9
sce.HMC9 <- sce[,colData(sce)$Sample=="HMC9"]
#Finding hvgs
dec.sce <- modelGeneVar(sce.HMC9)
fit.sce <- metadata(dec.sce)
plot(fit.sce$mean, fit.sce$var, xlab="Mean of log-expression", ylab="Variance of log-expression")
curve(fit.sce$trend(x), col="dodgerblue", add=TRUE, lwd=2)
chosen <- getTopHVGs(dec.sce, prop=0.1)
str(chosen)
sce.hvg.sub <- sce.HMC9[chosen,]
dim(sce.hvg.sub)
altExp(sce.hvg.sub, "original") <-sce.HMC9
altExpNames(sce.hvg.sub)
sce.hvg.sub

#Run PCA
set.seed(100)
sce.hvg.sub <- runPCA(sce.hvg.sub)

#Plot PCA
PCA <- as.data.frame(reducedDim(sce.hvg.sub, "PCA"))
colData(sce.hvg.sub)$PCA1 <- PCA$PC1
colData(sce.hvg.sub)$PCA2 <- PCA$PC2
PCAs.ident <- ggcells(sce.hvg.sub, aes(x=PCA1, y=PCA2, color=Identity))+
    geom_point() +
    xlab("PC1") +
    ylab("PC2") +
    ggtitle("PCAs") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
    scale_color_manual(values=setNames(colours.classes, classes))
PCAs.ident 
pdf("../../data/Supp_Fig_8/PCA_1_2_HMC9_Identity.pdf", width=6, height=4)
PCAs.ident
dev.off()
```
*Saving session info*
```{r}
sessionInfo()
```