---
title: "Supp_Fig_13"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Make sure before running this script you have run "Fig_3c.Rmd" so that you have generated the "sce_immunesub.rds" file.

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(knitr)

#libraries for reformatting data for bulk DEG analysis
library(SingleCellExperiment)
library(edgeR)
library(EnhancedVolcano)

#specific for GO term analysis
library(topGO)
#library(org.Hs.eg.db)

#specific for dendogram+heatmap
library(cowplot)
library(viridis)
library(RColorBrewer)
library(pheatmap)
library(gridExtra)
library(gridGraphics)

#load data
sce.sub <- readRDS("../../data/sce_immunesub.rds")
```
*Check sce*
```{r}
sce.sub
```
***Supp_Fig_13***
*Heatmap displaying the expression of genes characteristic of different immune cell subpopulations across both lactation and non-lactation associated mammary cells (LMC and NMCs).*
```{r}
m.imm <- as.matrix(logcounts(sce.sub))
pD.sub <- as.data.frame(colData(sce.sub))

#Load genes
genes <- c("CD68","FCER1G","CSF1R","ALCAM","CD86","HLA-DRB1","ICAM1","CD14","FCGR3A","ITGAX","ITGAM","ANPEP","C1QB","MMP9","CD163","CD33","MSR1","IL3RA","LILRA4","NT5E","PDGFRB","VCAM1","MPO","CD209","CD34","TNFRSF4","CD4","IL7R","TIGIT","LAG3","PDCD1","CTLA4","CD3D","BCL11B","LCK","KLRC1","FYN","CD8A","CD40LG","FOXP3","BATF","GZMA","GZMB","GZMH","GNLY","KLRD1","XCL1","NCAM1","CXCL13","CD200","JCHAIN","CD19","CD27","CD79A","MS4A1","FCGR3B","CEACAM8")

#Load colours
my_colour_sub = list(
    Sample=c(HMC1 = "salmon1", HMC2 = "peachpuff2", HMC3 = "burlywood", HMC4 = "#F8766D", HMC2B="peachpuff3", HMC5="salmon2", HMC6="pink1", HMC7="indianred2", HMC8="indianred4", HMC9= "hotpink2", RB1 = "steelblue3", RB2 = "slategray3", RB3 = "paleturquoise3", RB4 = "#00BFC4", RB5 = "turquoise", RB6= "turquoise4", RB7="lightblue1", RB8="lightskyblue2"),
    GraphClusters.sub=c(C1 = "lightgoldenrod3", C2 = "lightgoldenrod4", C3 = "olivedrab", C4 = "palegreen3", C5 = "lemonchiffon2", C6 = "gold4", C7 = "#99A800", C8="gold", C9="darkolivegreen",C10 ="lightsalmon3" , C11 ="yellow4" , C12 = "tan1", C13 = "palegreen4", C14 = "palegreen3", C15 = "orange1"))

#Heatmap for graphclusters.sub
subsP <- group_by(pD.sub, GraphClusters.sub) %>%
        mutate(GraphClusters.sub=forcats::fct_relevel(GraphClusters.sub,c("C9","C11","C12","C8","C2","C7","C6","C4","C3","C10","C5","C1"))) %>%
        arrange(GraphClusters.sub, Sample)

subsP <- as.data.frame(subsP)

all(genes %in% rownames(m.imm))
mheat <- m.imm[genes,as.character(subsP$Barcode)]

mheat <- mheat/apply(mheat,1,max) # Scale to 0-1 for visualization

#Set data for heat-map column name annotations
annoCol <- data.frame(Sample=subsP$Sample, GraphClusters.sub=subsP$GraphClusters.sub)
row.names(annoCol) <- subsP$Barcode
annoCol %>% arrange(GraphClusters.sub, Sample)

y <-  pheatmap(mheat,
	 cluster_rows=TRUE,
	 cluster_cols=FALSE,
         show_rownames=TRUE,
         show_colnames=FALSE,
         annotation_legend=TRUE,
	       annotation_col=annoCol,
         annotation_colors=my_colour_sub,
	      # gaps_col=c(1200, 1400),
	       fontsize=12)

ggsave(filename="../../data/Supp_Fig_13/Supp_Fig_13.pdf",y[[4]],width=8,height=7)
```
*Saving session info*
```{r}
sessionInfo()
```