---
title: "Supp_Fig_11"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scran)
library(scater)
library(ggplot2)
library(knitr)

#libraries for reformatting data for bulk DEG analysis
library(SingleCellExperiment)
library(edgeR)

##load data
sce <- readRDS("../../data/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))
```
*Check that the data is loaded correctly*
```{r}
sce
```

***Supp Fig11***
*MA plot of genes differentially expressed genes between luminal clusters LC1 and LC2 per milk cell sample.*
First we must generate a psuedobulk score for each sample by each cluster, then, after normalising compare the expression for each gene between the two groups and display on an MA plot.

*Running DEG analysis*
Generating psuedo-bulk samples from luminal HMCs
```{r}
#Create pseudo-bulk samples from the Samples subsetted according to GraphClusters
summed <- aggregateAcrossCells(sce,
                               ids=DataFrame(label=sce$Identity,
                                             sample=sce$Sample))
summed
table(summed$sample,summed$label)
```
*DEG analysis on subsetted HMC luminal cells from LC1 and LC2 *
```{r}
y<- DGEList(counts(summed), samples = colData(summed))
y

#pre-processing
discarded <- isOutlier(y$samples$lib.size, log=TRUE, type="lower")
y <- y[,!discarded]
summary(discarded)

#remove lowly expressed genes
keep <- filterByExpr(y, group=summed$Identity)
y <- y[keep,]
summary(keep)

#correct for compositional bias
y <- calcNormFactors(y)
y$samples
```
*Subset and generate MA plot per sample*
*LMC1*
```{r}
#HMC1
y.HMC1 <- y[,y$samples$sample=="HMC1"]
is.LC1 <- y.HMC1$samples$label=="LC1"

LC1<- calculateAverage((y.HMC1[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC1[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC1_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC1_LC1vsLC2_logFConly.csv")

HM <- data.frame(Genes = names(logFC), logFC.HMC1 = logFC)
```
*LMC2*
```{r}
#HMC2
y.HMC2 <- y[,y$samples$sample=="HMC2"]
is.LC1 <- y.HMC2$samples$label=="LC1"

LC1<- calculateAverage((y.HMC2[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC2[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC2_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC2 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC2_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC2B*
```{r}
#HMC2B
y.HMC2B <- y[,y$samples$sample=="HMC2B"]
is.LC1 <- y.HMC2B$samples$label=="LC1"

LC1<- calculateAverage((y.HMC2B[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC2B[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC2B_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC2B = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC2B_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC3*
```{r}
#HMC3
y.HMC3 <- y[,y$samples$sample=="HMC3"]
is.LC1 <- y.HMC3$samples$label=="LC1"

LC1<- calculateAverage((y.HMC3[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC3[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC3_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC3 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC3_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC4*
```{r}
#HMC4
y.HMC4 <- y[,y$samples$sample=="HMC4"]
is.LC1 <- y.HMC4$samples$label=="LC1"

LC1<- calculateAverage((y.HMC4[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC4[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC4_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC4 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC4_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC5*
```{r}
#HMC5
y.HMC5 <- y[,y$samples$sample=="HMC5"]
is.LC1 <- y.HMC5$samples$label=="LC1"

LC1<- calculateAverage((y.HMC5[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC5[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC5_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC5 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC5_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*HMC6*
```{r}
#HMC6
y.HMC6 <- y[,y$samples$sample=="HMC6"]
is.LC1 <- y.HMC6$samples$label=="LC1"

LC1<- calculateAverage((y.HMC6[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC6[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC6_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC6 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC6_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC7*
```{r}
#HMC7
y.HMC7 <- y[,y$samples$sample=="HMC7"]
is.LC1 <- y.HMC7$samples$label=="LC1"

LC1<- calculateAverage((y.HMC7[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC7[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC7_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC7 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC7_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC8*
```{r}
#HMC8
y.HMC8 <- y[,y$samples$sample=="HMC8"]
is.LC1 <- y.HMC8$samples$label=="LC1"

LC1<- calculateAverage((y.HMC8[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC8[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC8_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC8 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC8_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
```
*LMC9*
```{r}
#HMC9
y.HMC9 <- y[,y$samples$sample=="HMC9"]
is.LC1 <- y.HMC9$samples$label=="LC1"

LC1<- calculateAverage((y.HMC9[,is.LC1])$counts)
LC2 <- calculateAverage((y.HMC9[,!is.LC1])$counts)
library(scattermore)
logged <- cpm(cbind(LC1, LC2), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
fplot <- data.frame("abundance"=abundance,
            "logFC"=logFC,
            "Name"=names(logFC))
fplot <- dplyr::arrange(fplot,desc(abs(logFC)))
top <- fplot[fplot$logFC > 0,][1:15,]
btm <- fplot[fplot$logFC < 0,][1:15,]
ma.plt <- ggplot(fplot, aes(x=abundance, y=logFC, colour = (logFC > 2 | logFC < -2))) +
    geom_point() +
  scale_colour_manual(name = 'logFC', values = setNames(c('red','black'),c(T, F))) +
    geom_text_repel(data=top, size=5, aes(label=Name), colour="black") +
    geom_text_repel(data=btm, size=5, aes(label=Name), colour="black") +
    ggtitle("LC1 vs. LC2") +
    ylab("logFC") +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
    geom_hline(yintercept=0, color="black") +
    geom_hline(yintercept=c(-1.5,1.5), linetype = "dashed", color="black") +
    xlab("Average logExpr")
ma.plt
pdf("../../data/Supp_Fig_3/HMC9_LC1vsLC2_MAplot.pdf", width=10, height=6)
ma.plt
dev.off()

logFC.df <- data.frame(Genes = names(logFC), logFC.HMC9 = logFC)
logFC.df <- logFC.df[order(-logFC.df$logFC),]
write.csv(logFC.df, file= "../../data/Supp_Fig_3/HMC9_LC1vsLC2_logFConly.csv")

HM <- merge(HM,logFC.df, by="Genes")
write.csv(HM, file= "../../data/Supp_Fig_3/AllHMC_LC1vsLC2_logFConly.csv")
```
*Saving session info*
```{r}
sessionInfo()
```