---
title: "Supp_Fig_10_c"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
Make sure before running this script you have run "Fig_3_c.Rmd" as the output from that analysis is used to run GO term analysis.

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(ggplot2)

#specific for GO term analysis
library(topGO)
library(org.Hs.eg.db)

#specific for dendogram+heatmap
library(cowplot)
library(viridis)
library(RColorBrewer)
library(pheatmap)
library(gridExtra)
library(gridGraphics)

#load data
DEGS <- read.csv("../../data/Supp_Tables/Supplementary_Table_1.csv",stringsAsFactors = FALSE)
```
*Selecting highly significant genes for GO term analysis*
Making data frame of significant genes (p< 10e-8) and stipulating the logFC threshold for both upregulated and downregulated genes
```{r}
#all significant genes
sig.genes <- subset(DEGs, DEGs$FDR < 10e-8)
dim(sig.genes) #3422

#number of upregulated genes -> taking top 10%
upgenes <- subset(sig.genes, sig.genes$logFC > 0.01)
dim(upgenes) #1782
nup <- 0.05*nrow(upgenes)
nup #90
#number of downregulated genes -> taking top 10%
downgenes <- subset(sig.genes, sig.genes$logFC < -0.01)
dim(downgenes) #1640
ndown <- 0.05*nrow(downgenes)
ndown #82
```
***Supp Fig 10ci***
*The top 10 biological process gene ontology pathways that were associated with genes significantly differentially expressed that were either upregulated in LC1*
```{r}
#subset data frame for upgenes
upgenes <- sig.genes[order(-sig.genes$logFC),]
upgenes <- upgenes[1:nup,]
summary(upgenes)
dim(upgenes)

#generating named vector "geneList" which 
geneList <- as.integer(rownames(res$table) %in% rownames(upgenes))
names(geneList) <- rownames(res$table)

topDiffGenes <- function(allScore) {
 return(allScore > 0.01)
 }
 x <- topDiffGenes(geneList)
 sum(x)

#Creating the "topGOdata" object with at least 10 genes associated with each GO term
GOdata <- new("topGOdata",
  description = "GO analysis of luminal cell clusters; HMCcluster1vs2",
  ontology = "BP",
  allGenes = geneList,
  geneSel = topDiffGenes,
  annot = annFUN.org,
  mapping = "org.Hs.eg",
  ID = "symbol",
  nodeSize = 10)
GOdata

#To get the significant genes
sg <- sigGenes(GOdata)
str(sg)
numSigGenes(GOdata)

#Perform enrichment test, using the algorithm="weight01" (rather than algorithm = "classic") we take the GO hierarchy in to consideration
resultFisher <- runTest(GOdata, algorithm = "weight01", statistic = "fisher")

allRes <- GenTable(GOdata, classicFisher = resultFisher, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 50, numChar=200)
allRes

#add significant genes to the table
allRes$genes <- sapply(allRes$GO.ID, function(x)
    {
      genes<-genesInTerm(GOdata, x)
      genes[[1]][genes[[1]] %in% sg] # myGenes is the queried gene list
    })
allRes$genes[which(allRes$topGO<0.0001)] # print those only with p-value < 0.05
allRes$genes <-vapply(allRes$genes, paste, collapse = ", ", character(1L))

#Finding genes of significant annotations
length(usedGO(GOdata))

#Adding GeneRatio to the table which is number of significant genes annotated to the particular term divided by all terms used in the GO term analysis + making sure classidFisher registers as numeric for colour scales and ensuring "Term" is recognised not as a character so the plots will display the variables in order
allRes$GeneRatio <- allRes$Significant/length(sigGenes(GOdata))
allRes$classicFisher <- as.numeric(allRes$classicFisher)
allRes$Term <- factor(allRes$Term, levels=allRes$Term)
write.csv(allRes, file= "../../data/Supp_Tables/Supplementary_Table_3.csv")

allRes <- allRes %>% mutate(row = row_number())

#Plotting GO terms as a bargraph, where x-axis is GeneRatio, colour is classicfisher p-value
q <- ggplot(allRes[1:10,],
      aes(reorder(Term,-row))) +
      ylab("GeneRatio") +
      xlab("Term") +
      geom_bar(aes(weight = GeneRatio, fill=classicFisher)) +
      coord_flip() +
      theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + scale_fill_gradientn(colours= c("orchid","plum1"))
q
ggsave(filename="../../data/RB_HMC_merged/Supp_Fig_10/Fig_10cii.pdf",width=6,height=3)
```
***Supp Fig 10cii***
*The top 10 biological process gene ontology pathways that were associated with genes significantly differentially expressed that were either upregulated in LC2 *
```{r}
downgenes <- sig.genes[order(sig.genes$logFC),]
downgenes <- downgenes[1:ndown,]
summary(downgenes)
dim(downgenes)

#generating named vector "geneList" which 
geneList <- as.integer(rownames(res$table) %in% rownames(downgenes))
names(geneList) <- rownames(res$table)

#setting all genes >0.01 (i.e. =1, part of our subset) as "significant"
topDiffGenes <- function(allScore) {
 return(allScore > 0.01)
 }
 x <- topDiffGenes(geneList)
 sum(x)

#Creating the "topGOdata" object
GOdata <- new("topGOdata",
  description = "GO analysis of luminal cell clusters; HMCcluster1vs2",
 # ontology = "MF",
  ontology = "BP",
  allGenes = geneList,
  geneSel = topDiffGenes,
  annot = annFUN.org,
  mapping = "org.Hs.eg",
  ID = "symbol",
  nodeSize = 10)

GOdata

#To get the significant genes
sg <- sigGenes(GOdata)
str(sg)
numSigGenes(GOdata)
#60 significant genes

#Perform enrichment test, using the algorithum="weight01" (rather than algorithm = "classic") we take the GO hierarchy in to consideration
resultFisher <- runTest(GOdata, algorithm = "weight01", statistic = "fisher")

allRes <- GenTable(GOdata, classicFisher = resultFisher, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 50, numChar=200)
allRes

#add significant genes to the table
allRes$genes <- sapply(allRes$GO.ID, function(x)
    {
      genes<-genesInTerm(GOdata, x)
      genes[[1]][genes[[1]] %in% sg] # myGenes is the queried gene list
    })
allRes$genes[which(allRes$topGO<0.0001)] # print those only with p-value < 0.05
allRes$genes <-vapply(allRes$genes, paste, collapse = ", ", character(1L))

#Finding genes of significant annotations
length(usedGO(GOdata))

#Adding GeneRatio to the table which is number of significant genes annotated to the particular term divided by all terms used in the GO term analysis + making sure classidFisher registers as numeric for colour scales and ensuring "Term" is recognised not as a character so the plots will display the variables in order
allRes$GeneRatio <- allRes$Significant/length(sigGenes(GOdata))
allRes$classicFisher <- as.numeric(allRes$classicFisher)
is.na(allRes$classicFisher) <- 0.0000000000001
allRes$Term <- factor(allRes$Term, levels=allRes$Term)
allRes
write.csv(allRes, file= "../../data/Supp_Tables/Supplementary_Table_2.csv")

allRes <- allRes %>% mutate(row = row_number())

q <- ggplot(allRes[1:10,],
      aes(reorder(Term,-row))) +
      ylab("GeneRatio") +
      xlab("Term") +
      geom_bar(aes(weight = GeneRatio, fill=classicFisher)) +
      coord_flip() +
      theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + scale_fill_gradientn(colours= c("slateblue2","lightgray"))
q
ggsave(filename="../../data/Supp_Fig_10/Fig_10ci.pdf",width=9,height=3)
```
*Saving session info*
```{r}
sessionInfo()
```