---
title: "Supp_Fig_3"
author: "Alecia-Jane Twigger"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

**Setting up**

*Load data and packages*
Firstly we must load the important libraries and then the data.
```{r, message=FALSE}
#load libraries
library(plyr)
library(dplyr)
library(scater)
library(RColorBrewer)
library(ggplot2)

##load data

#----------#Pre-filtered data#----------#

#---#Batch1#---#
sce1 <- readRDS("../../data/Batch1/sce.rds")
rownames(rowData(sce)) <- uniquifyFeatureNames(rowData(sce)$ID,rowData(sce)$Symbol)
rownames(counts(sce)) <- rownames(rowData(sce))

#---#Batch2#---#
sce2 <- readRDS("../../data/Batch2/sce.rds")
rownames(rowData(sce)) <- uniquifyFeatureNames(rowData(sce)$ID,rowData(sce)$Symbol)
rownames(counts(sce)) <- rownames(rowData(sce))

#---#Batch3#---#
dataList <- readRDS("../../data/Batch3/ExpressionList.rds")
pD <- dataList$phenoData
fD <- dataList$featureData
m <- as(dataList$counts,"dgCMatrix")
rm(dataList)

#Generate a sce object
sce3 <- SingleCellExperiment(list(counts=as.matrix(m)),
    colData=DataFrame(pD),
    rowData=DataFrame(fD))

#Add lane info
colData(sce3)$Lane <- colData(sce)$Sample

#Mapping samples so they will fit with all merged data
sample.old <- c("HMC","RB1", "RB2","RB3")
sample.new <- c("HMC9","RB8","RB8","RB8")
colData(sce3)$Sample <- mapvalues(colData(sce3)$Sample,sample.old, sample.new)

state.old <- c("HMC","RB")
state.new <- c("HM","RB")
colData(sce3)$State <- mapvalues(colData(sce3)$State,state.old, state.new)


#----------#Post-filtered data#----------#
sce <- readRDS("../../data/sce_all_nospike_2.rds")
ClustCol <- as.data.frame(read.csv("../../data/Clusters_noSpike_2.csv",stringsAsFactors = FALSE))

pD <- as.data.frame(colData(sce))

#order samples in a particular order
pD.ord <- group_by(pD, Sample) %>%
        ungroup(Sample) %>%
  mutate(Sample=forcats::fct_relevel(Sample, c("HMC1", "HMC2", "HMC3","HMC4", "HMC9", "HMC2B","HMC5","HMC6","HMC7","HMC8","RB1", "RB2","RB3","RB4","RB8","RB5","RB6","RB7"))) %>%
        arrange(Sample)
```
*Check that the pre-filtered data is loaded correctly*
```{r}
sce1
sce2
sce3
```
*Check that the post-filtered data is loaded correctly*
```{r}
sce
```
*Adding quickPerCellQC*
```{r}
sce1 <- addPerCellQC(sce1, subsets=list(Mito=rowData(sce1)$Mito))
sce2 <- addPerCellQC(sce2, subsets=list(Mito=rowData(sce2)$Mito))
sce3 <- addPerCellQC(sce3, subsets=list(Mito=rowData(sce3)$Mito))
```
*Merging colData*
```{r}
sample.data <- rbind(colData(sce1),colData(sce2),colData(sce3))
pD.all <- as.data.frame(sample.data)
```
*Adding colours*
```{r}
sample.set <- c("HMC1", "HMC2", "HMC3","HMC4", "HMC2B","HMC5","HMC6","HMC7","HMC8", "HMC9","RB1", "RB2","RB3","RB4","RB5","RB6","RB7", "RB8")
colour.names <- c("salmon1", "peachpuff2", "burlywood", "#F8766D","peachpuff3", "salmon2", "pink1",  "indianred2", "indianred4", "hotpink2", "steelblue3", "slategray3", "paleturquoise3","#00BFC4","turquoise", "turquoise4", "lightblue1", "lightskyblue2")

#State colours
state <- c("RB", "HM")
state.colours <- c("#00BFC4","#F8766D")
```
***Pre-filtered plots***
**Supplementary Figure 3ai**
*Violin plot of the unique molecular identifiers (UMIs) expressed per cell per sample before filtering and normalisation*
```{r}
LibrarySize <- ggplot(pD.all, aes(x=Sample,y=sum, fill=Sample)) +
    geom_violin(draw_quantiles=0.5)+
    scale_y_log10() +
    ylab("Total number of molecules") +
    ggtitle("Number of Molecules") +
    geom_hline(yintercept=1000, linetype="dashed", color="red") +
    theme_bw() + scale_fill_manual(values=setNames(colour.names, sample.set)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
LibrarySize
pdf("../../data/Supp_Fig_3/Supp_Fig_3ai.pdf", width=12, height=8)
LibrarySize
dev.off()
```
**Supplementary Figure 3bi**
*Violin plot of the genes expressed per cell per sample before filtering and normalisation*
```{r}
#Violin plot of genes/cell per sample
genesDetected <- ggplot(pD.all, aes(x=Sample, y=detected, fill=Sample)) +
  geom_violin(draw_quantiles=0.5) +
  scale_y_log10() +
  ylab("Total number of genes detected") +
  ggtitle("Number of genes detected") +
  theme_bw() + scale_fill_manual(values=setNames(colour.names, sample.set)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
genesDetected
pdf("../../data/Supp_Fig_3/Supp_Fig_3bi.pdf", width=12, height=8)
genesDetected
dev.off()
```
**Supplementary Figure 3ci-ii**
*Percentage of mitochondrial genes compared with number of UMI per cell before filtering/normalisation coloured either by i) sample or ii) state.*
```{r}
#sce1
pD.sce1 <- as.data.frame(colData(sce1))
mad(pD.sce1$subsets_Mito_percent) #7.62695
mean(pD.sce1$subsets_Mito_percent) #8.088972
#cutoff=15.71592

#sce2
pD.sce2 <- as.data.frame(colData(sce2))
mad(pD.sce2$subsets_Mito_percent) #3.037994
mean(pD.sce2$subsets_Mito_percent) #6.32055
#cutoff=9.358544

#sce3
pD.sce3 <- as.data.frame(colData(sce3))
mad(pD.sce3$subsets_Mito_percent) #4.730828
mean(pD.sce3$subsets_Mito_percent) #7.75325
#cutoff=12.48408

#all together
mad(pD.all$subsets_Mito_percent) #5.547367
mean(pD.all$subsets_Mito_percent) #7.408134
#cutoff=12.9555

#Colour by Sample
MitoUMI <- ggplot(pD.all, aes(x=subsets_Mito_percent, y=sum))+
    geom_point(aes(shape=factor(State), colour=Sample), alpha=1/2) +
    xlab("Percentage of mitochondrial reads") +
    geom_vline(xintercept=12.95, linetype="dashed", color="red") +
    ggtitle("%Mito vs. UMIs") + scale_color_manual(values=setNames(colour.names, sample.set)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
MitoUMI
pdf("../../data/Supp_Fig_3/Supp_Fig_3ci.pdf", width=12, height=8)
MitoUMI
dev.off()

#Colour by State
MitoUMI <- ggplot(pD.all, aes(x=subsets_Mito_percent, y=sum))+
    geom_point(aes(shape=factor(State), colour=State), alpha=1/2) +
    xlab("Percentage of mitochondrial reads") +
    geom_vline(xintercept=12.95, linetype="dashed", color="red") +
    ggtitle("%Mito vs. UMIs") + scale_color_manual(values=setNames(state.colours, state)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
MitoUMI
pdf("../../data/Supp_Fig_3/Supp_Fig_3cii.pdf", width=12, height=8)
MitoUMI
dev.off()
```
***Pre-filtered plots***
**Supplementary Fig 3aii**
*Violin plots of the number of unique molecular identifiers (UMIs) expressed per cell per sample after filtering and normalisation.*
```{r}
LibrarySize <- ggplot(pD.ord, aes(x=Sample,y=sum, fill=Sample)) +
    geom_violin(draw_quantiles=0.5)+
    scale_y_log10() +
    ylab("Total number of molecules") +
    ggtitle("Number of Molecules") +
    theme_bw() + scale_fill_manual(values=setNames(colour.names, sample.set)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
LibrarySize
pdf("../../data/Supp_Fig_3/Supp_Fig_3aii.pdf", width=12, height=8)
LibrarySize
dev.off()
```
***Supplementary Fig 3bii***
*Violin plots of the number of genes expressed per cell per sample after filtering and normalisation.*
```{r}
genesDetected <- ggplot(pD.ord, aes(x=Sample, y=detected, fill=Sample)) +
  geom_violin(draw_quantiles=0.5) +
  scale_y_log10() +
  ylab("Total number of genes detected") +
  ggtitle("Number of genes detected") +
  theme_bw() + scale_fill_manual(values=setNames(colour.names, sample.set)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
genesDetected
pdf("../../data/Supp_Fig_3/Supp_Fig_3bii.pdf", width=12, height=8)
genesDetected
dev.off()
```
**Supplementary Fig 3ciii-v**
*Percentage of mitochondrial genes compared with number of genes per cell where cells are coloured by batch and shaped by mammary cell state (either lactating or non-lactating mammary cells LMC or NMC).*
```{r}
#Supplementary Fig 3ciii
MitoUMI <- ggcells(sce, aes(x=subsets_Mito_percent, y=sum))+
    geom_point(aes(shape=factor(State), colour=Sample), alpha=1/2) +
    xlab("Percentage of mitochondrial reads") +
    ggtitle("%Mito vs. UMIs") + scale_color_manual(values=setNames(colour.names, sample.set)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
MitoUMI
pdf("../../data/Supp_Fig_3/Supp_Fig_3ciii.pdf", width=12, height=8)

#Supplementary Fig 3civ
sce.HM <- sce[,colData(sce)$State=="HM"]
MitoUMI <- ggcells(sce.HM, aes(x=subsets_Mito_percent, y=sum))+
    geom_point(aes(shape=factor(State), colour=State), alpha=1/2) +
    xlab("Percentage of mitochondrial reads") +
    ggtitle("%Mito vs. UMIs") + scale_color_manual(values=setNames(state.colours, state)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
MitoUMI
pdf("../../data/Supp_Fig_3/Supp_Fig_3civ.pdf", width=12, height=8)

#Supplementary Fig 3cv
sce.RB <- sce[,colData(sce)$State=="RB"]
MitoUMI <- ggcells(sce.RB, aes(x=subsets_Mito_percent, y=sum))+
    geom_point(aes(shape=factor(State), colour=State), alpha=1/2) +
    xlab("Percentage of mitochondrial reads") +
    ggtitle("%Mito vs. UMIs") + scale_color_manual(values=setNames(state.colours, state)) +
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
MitoUMI
pdf("../../data/Supp_Fig_3/Supp_Fig_3cv.pdf", width=12, height=8)
```
*Saving session info*
```{r}
sessionInfo()
```